\begin{abstract}
We consider multi-objective reinforcement learning problems where objectives come from an identical family---such as the class of reachability objectives---and may appear or disappear at runtime.
\GS{Discounted cumulative reward is too broad of a class (see comment in intro) just mention reachability.}
Our goal is to design adaptive policies that can efficiently adjust their behaviors as the set of active objectives changes.
We propose a modular framework where each objective is supported by a selfish local policy, and coordination is achieved through a novel \emph{auction}-based mechanism:
policies bid for the right to execute their actions, with bids reflecting the urgency of the current state.
The highest bidder selects the action, enabling a dynamic and interpretable trade-off among objectives.
To succeed, each policy must not only optimize its own objective, but also reason about the presence of other goals and learn to produce calibrated bids that reflect relative priority.
When objectives change, the system adapts by simply adding or removing the corresponding policies.
Moreover, as objectives arise from the same family, identical copies of a parameterized policy can be deployed.
In our implementation, the policies are trained concurrently using proximal policy optimization (PPO).
We evaluate on Atari Assault and a gridworld-based path-planning task with dynamic targets.
Our method achieves substantially better performance than monolithic policies trained with PPO. 
\end{abstract}